# Control Flow
## Expession
can be simple expressions or compound expressions consisting of simple expression separated by semicolons or newlines and grouped within braces
```{r}
x <- {10 ; 20} # return the final expression value like rust
x
z = {x = 10; y = x^2; x + y}
x 
y 
z
```
## If The Else Statement
```{r}
# c++ way
x <- 10
if (x > 0) y = sqrt(x) else y = -sqrt(x)
y

# rust or lisp or algol way
y = if (x > 0) sqrt(x) else -sqrt(x)
y
```
### Combining Logical Conditions
`&&` and `||` operate evaluate just enough of their arguments
```{r}
TRUE && print(TRUE)
## both arguments evaluated
TRUE || print(TRUE)
## only evaluate the first, because it is enough to know the whole results
```
## Loops
`for` loop
```{r}
s <- 0
x <- 1:10
for(i in 1:length(x)) s = s + x[i]
s
s = 0
for(elt in x) s = s + elt
s
```
`next` is R's continue
```{r}
x <- 1:10
s <- 0
for(elt in x) {
    if (elt < 5) next
    print(elt)
    s = s + elt 
}
s
```
`while` loop
```{r}
threshold <- 100
n <- 0
s <- 0
while (s < threshold) {
    n = n + 1
    s = s + n
}
c(n, s)
```

```{r}
# gucci gang
dfs = function(remain, temp_str, lst){
    if (remain == 0){
        lst <- append(lst, list(temp_str))
        return(lst)
    }

    if (remain >= 1) {
        next_str <- if (temp_str == "") "gang" else
            paste(temp_str, "gang", sep = "")
        lst <- dfs(remain - 1, next_str, lst);
    }

    if (remain >= 2) {
        next_str <- if (temp_str == "") "gucci" else
            paste(temp_str, "gucci", sep = "");
        lst <- dfs(remain - 2, next_str, lst)
    }

    lst
}

gucci.gang = function(x){
    if (mode(x) != 'numeric') {
        print("error");
        return(NULL);
    }

    lst <- list();
    y <- '';
    lst <- dfs(x, y, lst);
    lst <- as.character(lst);
    lst
}

gucci.gang(3)
```

`repeat` loop with break
```{r}
threshold <- 100
n <- 0
s <- 0
repeat {
    if (s > threshold) break
    n = n + 1
    s = s + n
}
c(n, s)
```

## The switch
### switch by index
if `expr` has a numeric value equal to n, then the nth of the code blocks

# Functions

# More Examples
## Binomial Probability
## Statistical Computations
## Rational Approximations

# Vectorized Computation
## Summation Evaluation
## Poker Hands
## Some Useful Functions
## Frequencies of Gap Lengths
