# Control Flow
## Expession
can be simple expressions or compound expressions consisting of simple expression separated by semicolons or newlines and grouped within braces
```{r}
x <- {10 ; 20} # return the final expression value like rust
x
z = {x = 10; y = x^2; x + y}
x 
y 
z
```
## If The Else Statement
```{r}
# c++ way
x <- 10
if (x > 0) y = sqrt(x) else y = -sqrt(x)
y

# rust or lisp or algol way
y = if (x > 0) sqrt(x) else -sqrt(x)
y
```
### Combining Logical Conditions
`&&` and `||` operate evaluate just enough of their arguments
```{r}
TRUE && print(TRUE)
## both arguments evaluated
TRUE || print(TRUE)
## only evaluate the first, because it is enough to know the whole results
```
## Loops
`for` loop
```{r}
s <- 0
x <- 1:10
for(i in 1:length(x)) s = s + x[i]
s
s = 0
for(elt in x) s = s + elt
s
```
`next` is R's continue
```{r}
x <- 1:10
s <- 0
for(elt in x) {
    if (elt < 5) next
    print(elt)
    s = s + elt 
}
s
```
`while` loop
```{r}
threshold <- 100
n <- 0
s <- 0
while (s < threshold) {
    n = n + 1
    s = s + n
}
c(n, s)
```

```{r}
# gucci gang
dfs = function(remain, temp_str, lst){
    if (remain == 0){
        lst <- append(lst, list(temp_str))
        return(lst)
    }

    if (remain >= 1) {
        next_str <- if (temp_str == "") "gang" else
            paste(temp_str, "gang", sep = "")
        lst <- dfs(remain - 1, next_str, lst);
    }

    if (remain >= 2) {
        next_str <- if (temp_str == "") "gucci" else
            paste(temp_str, "gucci", sep = "");
        lst <- dfs(remain - 2, next_str, lst)
    }

    lst
}

gucci.gang = function(x){
    if (mode(x) != 'numeric') {
        print("error");
        return(NULL);
    }

    lst <- list();
    y <- '';
    lst <- dfs(x, y, lst);
    lst <- as.character(lst);
    lst
}

gucci.gang(3)
```

`repeat` loop with break
```{r}
threshold <- 100
n <- 0
s <- 0
repeat {
    if (s > threshold) break
    n = n + 1
    s = s + n
}
c(n, s)
```

## The switch
### switch by index
if `expr` has a numeric value equal to n, then the nth of the code blocks is executed.
```{r}
k <- 5
switch(k, "I", "II", "III", "IV", "V", "VI")
```
if `n` falls outside, return `NULL`   
switch be tag
```{r}
x = rnorm(10) # create 5 number from random normal distribution
loc.est = "mean"
y <- switch(loc.est,
    median = median(x),
    mean = mean(x)
)
z <- mean(x)
z == y
```

if no matching tag:   
1. has an untagged code block, auto match    
2. has no untagged code block, return `NULL`

```{r}
3 == switch("c",
    a = 1, 
    3,
    b = 2
)

r <- switch("C", a = 1, b = 2)
mode(r)
```
if code blocks is empty, match the nearest tag, in this case "a" and "b" both return "gotcha"
```{r}
"gotcha" == switch("a",
    a = ,
    b = "gotcha",
    c = "missed"
)
```


# Functions
```{r}
square = function(x){
    x * x
}

square(1:10) # * is element wise

sumsq = function(x){
    sum(square(x))
}
sumsq(1:10)

sumsq = function(x){
    sum(square(x - mean(x)))
}
sumsq(1:10)

hypot = function(a, b){
    z <- sqrt(a ^ 2 + b ^ 2)
    return(z)
}
hypot(3, 4)
```
iterative procedure for computing square roots for positive numbers
```{r}
square.root = function(x, iterator = 10, start = 1){
    r.new = start
    r.old = 0
    while (iterator != 0){
        if (all(r.new == r.old)) break
        r.old = r.new
        r.new = (r.new + x / r.new) / 2
        iterator = iterator - 1
    }
    r.new
}
square.root(2, 10)
square.root(1:10)
```
## Binomial Probability
```{r}
# suppose a coin is tossed 100 times, the coin probebility is 50%, let's work out the probability of getting between 45 and 55
d <- dbinom(45:55, 100, 1/2)
d
sum(d)
# the chance between 40 and 60 in 100 tossed
round(sum(dbinom(40:60, 100, 0.5)), digits = 4)
# the chance between 35 and 65 heads in 100 tossed
round(sum(dbinom(35:65, 100, 0.5)), digits = 4)
```

### Coin Tossing Experiment
`runif(x)` generate x numbers between 0 and 1
```{r}
runif(10) < 0.5
sum(runif(10) < 0.5) # return logical vector coercion to numeric vector then sum, TRUE := 1, FALSE := 0
```

Coin Tossing: get the number of heads in 100 tosses, let's repoeat the process 10000 times and see wht proportion of values fall between 45 and 55
```{r}
nheads = numeric(10000)
nheads
for(i in 1:10000){
    nheads[i] = sum(runif(100) < 0.5)
}
mean(45 <= nheads & nheads <= 55)
```
package the experiment   
through the coin in ntosses round, has the phead, for a round like this repeat nreps times
```{r}
cointoss = function(ntosses, pheads, nreps = 1){
    nheads = numeric(nreps)
    for(i in 1:nreps){
        #0 -------- pheads ---------------- 1
        #   正面区域          反面区域
        nheads[i] = sum(runif(ntosses) < pheads)
    }

    nheads
}
```
## Statistical Computations
$n$ is sample size, $\overline{y}$ is the mean, $s$ is standard deviation
$$
    t = \frac{\overline{y_1} - \overline{y_2}}{s_p \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}} \
$$
$s_e$ is standard error
$$
    s_e = s_p \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}
$$
$s_p$ is pooled estimate of the standard deviation(合并标准差估计)
$$
    s_p^2 = \frac{(n_1 - 1)s_1^2 + (n_2 - 1)s_2^2}{n_1 + n_2 - 2} \\

    t = \frac{\overline{y_1} - \overline{y_2}}{s_e}
$$
`var(x)`：方差 $s^2$
`sd(x)`：标准差 $s$
```{r}
diff.mean = function(y1, y2) {
    mean(y1) - mean(y2)
}

sp = function(y1, y2){
    n1 = length(y1)
    n2 = length(y2)
    sqrt(((n1 - 1) * var(y1) + (n2 - 1) * var(y2)) / (n1 + n2 - 2))
}

pooled.se = function(y1, y2) {
    n1 = length(y1)
    n2 = length(y2)

    sp(y1, y2) * sqrt(1 / n1 + 1 / n2)
}

t.val = function(y1, y2){
    diff.mean(y1, y2) / pooled.se(y1, y2)
}

ttest = function(y1, y2){
    n1 = length(y1)
    n2 = length(y2)
    tval = t.val(y1, y2)
    pval = 2 * pt(-abs(tval), n1 + n2 - 2)
    list(t = tval, df = n1 + n2 -2, pval = pval)
    # df = degree of freedom
    # t = mean / 差距随机波动多少
    # pval = 观察到当前这么大或者更大的差距,概率为pval%
}

y1 = rnorm(10)
y2 = rnorm(10) + 2
y1
y2
ttest(y1, y2)
```

## Rational Approximations
```{r}
cf.expand = function(x, n = 5) {
    cf = numeric(n)
    for(i in 1:n){
        cf[i] = round(x)
        delta = x - cf[i]
        x = 1 / delta
    }
    cf
}
cf.expand(pi, 7)
```
continued fraction can be reduced to a simple ratio as follows
Start with $ N = c_n $ and $ D = 1$. Then for $ i = n-1\ldots1$ set
$$
    T = N \
    N = N * c_i + D
    D = T
$$

```{r}
rat.approx = function(x, n){
    cf = cf.expand(x, n)
    num = cf[n]
    den = 1
    if (n > 1){
        for (j in (n - 1) : 1){
            tmp = num
            num = cf[j] * tmp + den
            den = tmp
        }
    }

    if (den > 0){
        c(num, den)
    } else {
        c(-num, -den)
    }
}
rat.approx(pi, 1)
print(pi, digits = 15)
```

# Vectorized Computation
use the *apply mechanism only if the result has the same shape and already requires additional function calls
```
apply(x, MARGIN, FUN) x: Input Data MARGIN: row manipulate FUN: Column manipulate
lapply(X, FUN) l -> list x: Input Data, FUN: the operation you want to apply
sapply(X, FUN) return list | vector , see the compiler optimization if can return vector
vapply(X, FUN, FUN.VALUE) FUN.VALUE is the return value type
mapply(FUN, input...) mulitple apply
```
```{r}
# Stirling numbers 
s1 = function(n, k){
    x = 0:k
    temp = sapply(x, function(input){
        (-1) ^ (k - input) * choose(k, input) * input ^ n
    })
    (1 / factorial(k)) * sum(temp)
}
s1(10, 3)

s2 = function(n, k){
    x = 0:k
    sum( (-1) ^ (k - x) * choose(k, x) * x ^ n) / factorial(k)
}
s2(10, 3) == s1(10, 3)
```
## Poker Hands
```{r}
x = c(2, 1, 3, 2, 1, 2, 3, 3, 3, 1, 1, 3, 1, 3, 2, 2, 1, 2, 2, 3)
hands = function(x) {
    number.of.hands = length(x) %/% 3
    i = seq_len(number.of.hands)

    a = x[3 * i - 2]
    b = x[3 * i - 1]
    c = x[3 * i]

    hand1 = a != b & a != c & b != c
    hand3 = a == b & b == c
    hand2 = !hand1 & !hand3

    c(
        hand1 = sum(hand1),
        hand2 = sum(hand2),
        hand3 = sum(hand3)
    )
}
hands(x)
system.time(hands(x))
```

## Some Useful Functions
```{r}
which(c(TRUE, FALSE, FALSE, TRUE)) # return the indices of TRUEs
which.max(c(6:10, 13:5)) # return the first max/min index
diff(c(1, 3, 6, 4)) # return consecutive difference
x <- rbinom(20, 10, 0.5) + 1
x
tabulate(x) # 统计每个正整数出现了多少次
```
## Frequencies of Gap Lengths
