# R Basic

## Use R as Calculator

```{r}
1 + 2
1 / 2
17 ^ 2
1 + 2 * 3
(1 + 2) * 3
```

## Operator
- \^ exponentiation

```{r}
2 ^ 3 # 8
```

- \+ \- unary plus and minus

```{r}
+3
-(-1)
```

- : sequence operator
```{r}
1 : 3 # 1..3
```
- %/% integer division 

```{r}
7 %/% 3
```

- %% remainder

```{r}
7 %% 3
```

- \* / multiplication, division

```{r}
2 * 3
12 / 13 # output is a double or float
```

- \+ \- additon, subtraction

```{r}
1 + 2
1 - 2
```

## Evalution Order
Originally evaluation of equal precedence is left to right ->   
<mark>*Exponentiation $ \^ $ is right to left <-* </mark>


```{r}
2 ^ 3 ^ 2 # 2 ^ (9)
(2 ^ 3) ^ 2 # 8 ^ 2
```

## R build-in Function
1. sqrt(x)
$$
  \sqrt{x}
$$
```{r}
sqrt(12)
```
2. log(x, base=e) deflaut base is $e$
$$
    \log_e{x}
$$

```{r}
log(exp(1))
```
3. log10(x)
$$
    \log_{10}{x}
$$

```{r}
log10(10)
```

4. sin(x)
$$
    \sin{x}
$$

```{r}
sin(pi / 2)
```

5. atan(x)
$$
    \theta = \arctan{\frac{y}{x}}
$$

```{r}
4 * atan(1)
```

## Assignment and use variable

```{r}
z <- 17 # preferred since it is unambiguous
# it is not z < -17
17 -> z
z = 17 # in interactive use "="
```

## Numeric Vectors
c stands for combined, work with the vectors of values
```{r}
x <- c(1, 2, 3, 4)
x # can be viewed simply typing the name of the vector
```

## Sequences
$n_1 : n_2$ generate the sequence of integers ranging from n_1 to n_2
```{r}
1 : 50
5 : -5
```

## Combining Vectors
$c$ can be used to combine both vectors and scalars into large vectors  
Reason: R stores scalar values like 10 as vectors {10,} of length of one, so that both arguements in the expression above are vectors
```{r}
x = c(1, 2, 3, 4)
c(x, 10)
c(x, x)
```

## Vector Arithmetic
Because everything is a vector, so it is natural to expect that the type of arithemetic manipulation carried out in the previous section to also work with more general vectors

```{r}
2 * x + 1
sqrt(x)
```

## Recycling Rule of Vectors

```{r}
c(1, 2, 3, 4) + c(1, 2)
# c(1, 2, 3, 4) + c(1, 2, 1, 2)
```

$$
    \begin{bmatrix}
        1 \\
        2 \\
        3 \\
        4
    \end{bmatrix} + 
    \begin{bmatrix}
       1 \\
       2
    \end{bmatrix} \xrightarrow{recyle}
    \begin{bmatrix}
        1 \\
        2 \\
        3 \\
        4
    \end{bmatrix} +
    \begin{bmatrix}
       1 \\
       2 \\
       1 \\
       2
    \end{bmatrix} \xrightarrow{add}
    \begin{bmatrix}
        2 \\
        4 \\
        4 \\
        6
    \end{bmatrix}
$$

## Binary Operations
`+` addition  
`-` subtranction    
`*` multiplication  
`/` division  
`^` rasing to a power  
`%%` modulo (remainder after division)    
`%/%` integer division

## The modulo and Integer Division Operators
The modulo operator is useful in the integer computations
```{r}
11 %% 3
1:10 %% 2
```
but it can be used with more genreal numbers
```{r}
13.5 %% 2
13.5 %/% 2
```

## Elements and subsets
```{r}
# x <- c(1, 2, 3, 4)
x[3] # return the third element
x[c(1, 3)] # return the first and the third element
x[1:3] # return range (1, 2, 3) element
```

## Negative Subscripts
extract all the elements from x except the first and third
```{r}
# x <- c(1, 2, 3, 4)
x[c(-1, -3)] # x[2, 4]
```
It is not possible to mix positive and negative subscripts

## Changing Vector Subsets
```{r}
y <- 1:10
y[4:6] = 0
y
```

## Infinity Value
```{r}
1 / 0 
```
`Inf` represents positive infinity, `-Inf` represents negative infinity
```{r}
-1 / 0
1 + Inf
10000 / Inf
```

## Not a Number NaN
`NaN` indicates that a numerical result is undefined
```{r}
0 / 0
Inf - Inf
sqrt(-1)
```

## Not Available NA
The value is indicated by `NA`
```{r}
1 + sin(NA)
```
`NA` is used for statistical ovservations where the value could not be recorded

## Build in Function 
```{r}
abs(-12) # return absolute value
sign(12) # if + -> 1 else if - -> -1 else 0
sqrt(4) # return the square root
exp(2) # return e ^ n
log(exp(1)) # return natual logarithms
log10(10 ^ 2) # return common logarithms
sin(pi / 2)
cos(pi / 2)
tan(pi / 2) 
# return the trigonometric functions
asin(0)
acos(0)
atan(0) # return inverse trigonometric functions
factorial(3) # return factorial coefficient
choose(4, 2) # choose 2 out of 4 with no put back
round(1.7245, digits = 2) # round to two decimal places
signif(1.7, digits = 1) # round to two significant figures
ceiling(3.5)
floor(3.5)
# rounding up and down
```

# statistical Tables
- `d` -> probability densities function
- `p` -> cumulative probability function
- `q` -> inverse cumulative distribution function
- `r` -> random gen (r*(n: repeat_time, size: eachtime try how many times, prob: the probability of succes))

1. `norm` Normal Distribution
2. `chisq` Chi-squared Distribution
3. `t` t Distribution
4. `F` F Distribution
5. `binom` Binomial Distribution
6. `pois` Poisson distribution
```
# 1. 标准正态分布：mean = 0, sd = 1
pnorm(q, mean = 0, sd = 1)

# 2. 卡方分布：必须提供自由度 df
pchisq(q, df, ncp = 0)

# 3. t 分布：必须提供自由度 df
pt(q, df)

# 4. F 分布：必须提供两个自由度
pf(q, df1, df2, ncp = 0)

# 5. 二项分布：试验次数和成功概率都必须提供
pbinom(q, size, prob), q 最多成功多少次

# 6. 泊松分布：必须提供均值/发生率 lambda
ppois(q, lambda)
```

```{r}
pnorm(2)
pbinom(44, 100, 0.5)
dpois(3, 5)
```

## Summary Function
```{r}
sum(1:100) # 1 + ... 100
prod(1:10) # 1 * .. 10
```
special case with `NA`: in any summary functions that has `NA` or `NaN` values, always return `NA` or `NaN`
```{r}
min(10, NaN)
min(10, NA)
max(11, NA)
max(12, NaN)
min(NaN, NA) # return NA(Not Available)
max(NaN, NA) # return NA(Not Available)
min(10, 20, NA, na.rm = TRUE)
```

## Cumulative Summaries
These functions doesn't have `na.rm` argument (NA remove)
```{r}
cumsum(1:10) # 1 + ...
cumprod(1:6) # 1 * ...
cummax(1:10) # max(1,n)...
cummin(1:10) # min(1,n)...
```

## Parallel Summary Functions
Still have recycling rules, and accpet `na.rm` argument(NA remove)
```{r}
pmin(c(1, 10), c(10, 2))
pmax(0, c(-1, 0, 1))
```

## Type of Vectors
```{r}
b <- c(TRUE, TRUE, FALSE, FALSE)
b[1:3]
b[1]
b
```

## Logical Conjunctions
```{r}
x <- c(1, 2, 3, 4)
x < 2 | x > 3 # [TRUE, FALSE, FALSE, TRUE]
```
`any` for if any element matches the conditions then return true;    
`all` for if all elements match the conditions then return true;
```{r}
any(x >= 4) # return all true
all(x > 3) # return false
```
vectorised operation uses `&&` or `|`;   
if condition uses `&&` or `||`;   
negation uses `!`

## Negation
```{r}
!(3 < 4)
TRUE & !FALSE
```

## Logic and NA values
x & y    
- FALSE & NA -> FALSE   
- TRUE & NA -> NA   
- NA & NA -> NA    
x | y     
- FALSE | NA -> NA   
- TRUE | NA -> TRUE   
- NA | NA -> NA     

## logical subsetting
extract subsets of vectors by using logical values
```{r}
x <- c(1, 2, 3, 4)
x[x > 1]
```

## conditional selection
```{r}
x <- 1:10
ifelse(x > 5, x, -x)
ifelse(x > 5, x, 0)
```

## Character Vectors
```{r}
s <- c("first", "second", "third")
s[1:2]
s[1] = "initial"
s
```

## String Manipulation
```{r}
s
nchar(s)
substring(s, first = 1, last = 2:3)
substring(s, c(1,1,1), c(2, 3, 2))
substring(s, c(1,1,1), 2:3)
```

## Pasting Strings Together
`paste` is for glue string, with sep argument
```{r}
paste("First", "Second", "Third")
paste("First", "Second", "Third", sep=": ")
paste("First", "Second", "Third", sep="")
paste0("First", "Second", "Third")
```
## Pasting Vectors
```{r}
s <- c("first", "second", "third")
paste(s, "element", sep = "-")
# An additional argument, collapse
paste(s, collapse = " -> ")
```

## Complex valued Vectors
```{r}
z <- 10 + 20i
z
sqrt(z)
```

## Checking vector mode
4 mode:   
1. logical
2. numeric
3. complex
4. character
```{r}
mode(1:10)
length(1:10)
z <- logical(5) + 1
mode(z)
logical(5)
numeric(5)
```

## Creating Vectors
```{r}
vector('numeric', 5)
vector('logical', 5)
numeric(0)
```

## Type Coercion
logical -> numeric     
numeric -> complex (without imaginary part)   
all mode -> characteric    

## Type Coericion Idioms
```{r}
x <- 1:10
sum(x > 5) # ifelse(x > 5, TRUE, FLASE) -> sum
# x > 5 -> {0, 0, 0, 0, 0, 1, 1, 1, 1, 1}
# cumsum() -> {0, 0, 0, 0, 1, 1, 1, 1, 1}
cumsum(x > 5)
```

## Explicit Coercion
非0就是true
```{r}
as.numeric("1") + as.numeric('2')
as.logical(1) & as.logical(-1)
as.logical("TRUE")   # TRUE
as.logical("FALSE")  # FALSE
as.logical("abc")    # NA
```

## Named Vectors
```{r}
v = 1:4
names(v) = c("A", "B", "C", "D")
v
names(v)
```
## Subsetting Using Names
```{r}
v = c(1, 2, 3, 4)
names(v) = c("A", "B", "C", "D")
v['A']
v[c('A', 'D')]
```

## Lists
a way fo storing things of different types in a single object
```{r}
lst <- list(10, "eleven", TRUE, c(1, 2))
lst
```
ask for single element fo a list, use `[[]]`
```{r}
lst[[1]]
```
ask for a subset, use `[]`
```{r}
lst[1]
```

## Extracting Named Elements
```{r}
lst = list(a = 10, b = "eleven", TRUE)
lst[['a']]
lst$a
```

## NULL object
NULL represent nothing, has a length of 0 and can be freely coerced to either a vector or list of length 0
```{r}
length(NULL)
as.numeric(NULL)
as.list(NULL)
```

## Creating Patterned Sequences
```{r}
1:10
10:1
1.5:7.5
1.5:7
```

## Use seq() for more general purpose
```{r}
seq(0, 1, by = 0.1)
seq(0, 1, length = 11)
seq(0, by = 0.1, length = 11)
seq(to = 1, by = 0.1, length = 11)
x <- 1:10
seq(along = x)
x = numeric(0)
seq(along = x)
1: length(x)
```

## repeat values
`rep(x1, x2)`:   
x1 : the values to be repeated
x2 : how many times to repeat
```{r}
rep(1:4, 3)
rep(1:4, c(2, 3, 2 ,4))
rep(c('A', 'B', 'C'), 3)
```
`each` argument
```{r}
rep(1:4, each=3)
rep(1:4, rep(3, 4))
```